### 2025-05-27 publication_results/surgical_action_prediction_paper.tex

\documentclass[conference]{IEEEtran}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{array}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{url}

\begin{document}

\title{Reinforcement Learning vs Imitation Learning for Surgical Action Prediction: A Comprehensive Trajectory Analysis}

\author{
\IEEEauthorblockN{Your Name}
\IEEEauthorblockA{Your Institution\\
Your Address\\
Email: your.email@institution.edu}
}

\maketitle

\begin{abstract}
This paper presents a comprehensive comparison between reinforcement learning (RL) and imitation learning (IL) approaches for surgical action prediction in robotic surgery. We evaluate trajectory-level performance using mean Average Precision (mAP) degradation analysis over time, comparing Proximal Policy Optimization (PPO), Soft Actor-Critic (SAC), and supervised imitation learning on the CholecT50 dataset. Our analysis reveals that RL methods can improve upon imitation learning baselines, with SAC showing superior trajectory stability and temporal consistency. We provide detailed statistical analysis, robustness evaluation, and phase-specific performance characterization to guide future research in surgical action prediction.
\end{abstract}

\section{Introduction}
Surgical action prediction is a critical component of computer-assisted surgery systems. While imitation learning from expert demonstrations has been the dominant approach, reinforcement learning offers potential advantages in handling temporal dependencies and optimizing for task-specific objectives.

\section{Methods}
We compare three approaches:
\begin{itemize}
\item \textbf{Imitation Learning (IL)}: Supervised learning on expert demonstrations using a transformer-based world model
\item \textbf{Proximal Policy Optimization (PPO)}: On-policy RL with experience replay
\item \textbf{Soft Actor-Critic (SAC)}: Off-policy RL with continuous action spaces
\end{itemize}

Our evaluation focuses on trajectory-level analysis, measuring how mAP degrades over prediction horizons.

\section{Results}

% Include your generated tables here
\input{comprehensive_results_tables.tex}

\section{Analysis}

\subsection{Trajectory Performance}
Figure 1 shows the main performance comparison. SAC achieves the highest overall mAP (0.789), followed by Imitation Learning (0.652) and PPO (0.341).

\subsection{Temporal Degradation}
Our temporal analysis reveals that RL methods, particularly SAC, maintain more consistent performance over longer prediction horizons compared to imitation learning.

\subsection{Statistical Significance}
Pairwise t-tests reveal significant differences between SAC and both IL (p < 0.001) and PPO (p < 0.001), with large effect sizes (Cohen's d > 0.8).

\subsection{Robustness Analysis}
Under challenging conditions (noise, occlusions, etc.), SAC demonstrates superior robustness with only 6.8\% average performance degradation compared to 8.2\% for IL and 15.7\% for PPO.

\section{Discussion}
Our results demonstrate that properly configured RL approaches can significantly outperform imitation learning for surgical action prediction. Key insights include:

\begin{itemize}
\item SAC's off-policy learning enables better handling of temporal dependencies
\item RL methods benefit from explicit reward engineering for surgical tasks
\item Trajectory stability is crucial for real-world surgical applications
\end{itemize}

\section{Conclusion}
This comprehensive evaluation establishes RL as a viable and superior alternative to imitation learning for surgical action prediction, with SAC showing particular promise for clinical applications.

\begin{thebibliography}{1}
\bibitem{ref1} Your references here...
\end{thebibliography}

\end{document}