### 2025-06-03 archive_and_fixes/publication_results/comprehensive_results_tables.tex
clean root dir

\begin{table*}[htbp]
\centering
\caption{Comprehensive Comparison: Reinforcement Learning vs Imitation Learning for Surgical Action Prediction}
\label{tab:main_results}
\begin{tabular}{lccccccc}
\toprule
Method & Mean mAP & Start mAP & End mAP & Degradation & Rel. Deg. & Trajectory Slope & Stability Rank \\
\midrule

Imitation Learning & 0.349 & 0.998 & 0.953 & -0.177 & 4.5\% & -0.0003 & 1 \\
Sac & 0.278 & 0.989 & 0.256 & -0.092 & 74.2\% & -0.0058 & 2 \\
Ppo & 0.277 & 0.988 & 0.396 & -0.110 & 60.0\% & -0.0045 & 3 \\

\bottomrule
\multicolumn{8}{l}{\footnotesize mAP = mean Average Precision; Rel. Deg. = Relative Degradation} \\
\multicolumn{8}{l}{\footnotesize Stability Rank: 1 = most stable, higher = less stable} \\
\end{tabular}
\end{table*}

\begin{table}[htbp]
\centering
\caption{Trajectory mAP Degradation Analysis Over Time}
\label{tab:map_degradation}
\begin{tabular}{lcccc}
\toprule
Method & Initial Performance & Final Performance & Absolute Degradation & Significance \\
\midrule

Sac & 0.989 ± 0.007 & 0.256 ± 0.040 & 0.733 & *** \\
Ppo & 0.988 ± 0.008 & 0.396 ± 0.024 & 0.592 & *** \\
Imitation Learning & 0.998 ± 0.004 & 0.953 ± 0.014 & 0.045 & * \\

\bottomrule
\multicolumn{5}{l}{\footnotesize *** p < 0.001, ** p < 0.01, * p < 0.05} \\
\end{tabular}
\end{table}

\begin{table}[htbp]
\centering
\caption{Statistical Significance Tests: Pairwise Method Comparisons}
\label{tab:statistical_tests}
\begin{tabular}{lcccc}
\toprule
Comparison & Mean Difference & t-statistic & p-value & Effect Size (Cohen's d) \\
\midrule

Imitation Learning vs Ppo & 0.071 & 4.300 & 0.000*** & 0.272 \\
Imitation Learning vs Sac & 0.071 & 4.269 & 0.000*** & 0.270 \\
Ppo vs Sac & -0.000 & -0.013 & 0.990 & -0.001 \\

\bottomrule
\multicolumn{5}{l}{\footnotesize *** p < 0.001, ** p < 0.01, * p < 0.05} \\
\end{tabular}
\end{table}

\begin{table}[htbp]
\centering
\caption{Detailed Method Performance Characteristics}
\label{tab:method_comparison}
\begin{tabular}{lcccc}
\toprule
Method & Consistency & Robustness & Learning Type & Best Use Case \\
\midrule

Imitation Learning & High & Medium & Supervised & Stable environments \\
Ppo & Medium & Low & RL (On-policy) & Exploration-heavy tasks \\
Sac & High & High & RL (Off-policy) & Complex dynamics \\

\bottomrule
\end{tabular}
\end{table}