### 2025-06-03 archive/corrected_evaluation_results/corrected_latex_table.tex
add archives

\begin{table}[htb]
\centering
\caption{Trajectory mAP Analysis: Single-step Inference with Cumulative Evaluation}
\label{tab:trajectory_map_results}
\begin{tabular}{lcccc}
\toprule
Method & Mean mAP & Std mAP & Degradation & Videos \\
\midrule
Imitation Learning & 0.979 & 0.018 & 0.037 & 5 \\
Sac & 0.331 & 0.109 & 0.360 & 5 \\
Ppo & 0.000 & 0.000 & 0.000 & 5 \\

\bottomrule
\end{tabular}
\begin{tablenotes}
\small
\item Inference: Single-step prediction (predict $t+1$ from $t$)
\item Evaluation: Cumulative mAP (evaluate predictions from start to timestep $t$)
\item Degradation: mAP difference between first and last timestep
\end{tablenotes}
\end{table}