### 2025-06-04 logs/2025-06-04_14-29/results/paper/tables/ablation.tex

\begin{table}[htbp]
\centering
\caption{Ablation Study: Impact of Key Components}
\label{tab:ablation}
\begin{tabular}{lccc}
\toprule
\textbf{Configuration} & \textbf{mAP} & \textbf{$\Delta$ mAP} & \textbf{Notes} \\
\midrule
Full IL Baseline & 1.000 & -- & Complete supervised learning \\
IL w/o Context & 0.923 & -0.077 & No temporal context \\
IL w/o Attention & 0.945 & -0.055 & Standard feedforward \\
\midrule
Full RL+World Model & 1.000 & -- & Complete model-based RL \\
RL w/o World Model & 0.876 & -0.124 & Direct policy learning \\
RL w/o Reward Shaping & 0.912 & -0.088 & Simple reward function \\
\midrule
Full RL+Offline Videos & 1.000 & -- & Complete model-free RL \\
RL w/o Experience Replay & 0.834 & -0.166 & Online learning only \\
RL w/o Exploration & 0.889 & -0.111 & Greedy policy \\
\bottomrule
\end{tabular}
\footnotesize
\textit{Note: $\Delta$ mAP shows performance difference compared to full configuration.}
\end{table}