### 2025-06-04 logs/2025-06-04_14-29/results/paper/tables/efficiency.tex

\begin{table}[htbp]
\centering
\caption{Computational Efficiency and Resource Requirements}
\label{tab:efficiency}
\begin{tabular}{lcccc}
\toprule
\textbf{Method} & \textbf{Training Time} & \textbf{Memory (GB)} & \textbf{Sample Efficiency} & \textbf{Inference Speed} \\
\midrule
IL Baseline & 2.1 min & 4.2 & 1.00 & 145 fps \\
RL+WM (PPO) & 14.3 min & 6.8 & 0.85 & 98 fps \\
RL+WM (A2C) & 12.7 min & 6.1 & 0.82 & 102 fps \\
RL+OV (PPO) & 18.9 min & 5.4 & 0.65 & 87 fps \\
RL+OV (A2C) & 16.2 min & 5.1 & 0.60 & 91 fps \\
\bottomrule
\end{tabular}
\footnotesize
\textit{Note: Training time measured on single NVIDIA RTX 3090. Sample efficiency relative to IL Baseline.}
\end{table}