### 2025-06-04 logs/2025-06-04_14-29/results/paper/tables/main_results.tex

\begin{table}[htbp]
\centering
\caption{Performance Comparison of Surgical Action Prediction Methods}
\label{tab:main_results}
\begin{tabular}{lcccc}
\toprule
\textbf{Method} & \textbf{mAP} & \textbf{Degradation} & \textbf{Stability} & \textbf{Rank} \\
\midrule
RL WorldModel PPO & 1.000 & 0.000 & -0.000 & 1 \\
RL OfflineVideos A2C & 1.000 & 0.000 & -0.000 & 2 \\
IL Baseline & 1.000 & 0.000 & -0.000 & 3 \\
RL WorldModel A2C & 1.000 & 0.000 & -0.000 & 4 \\
RL OfflineVideos PPO & 0.960 & 0.010 & -0.010 & 5 \\

\bottomrule
\end{tabular}
\footnotesize
\textit{Note: mAP = Mean Average Precision, Degradation = Performance loss over planning horizon, Stability = Negative degradation score.}
\end{table}