### 2025-06-04 logs/2025-06-04_14-29/results/paper/tables/significance.tex

\begin{table}[htbp]
\centering
\caption{Statistical Significance Test Results (p-values)}
\label{tab:significance}
\begin{tabular}{lccccc}
\toprule
\textbf{Method} & \textbf{IL} & \textbf{WM-PPO} & \textbf{WM-A2C} & \textbf{OV-PPO} & \textbf{OV-A2C} \\
\midrule
IL Baseline & -- & 0.182 & 0.165 & 0.023* & 0.187 \\
RL+WM (PPO) & 0.182 & -- & 0.891 & 0.019* & 0.245 \\
RL+WM (A2C) & 0.165 & 0.891 & -- & 0.017* & 0.221 \\
RL+OV (PPO) & 0.023* & 0.019* & 0.017* & -- & 0.012* \\
RL+OV (A2C) & 0.187 & 0.245 & 0.221 & 0.012* & -- \\
\bottomrule
\end{tabular}
\footnotesize
\textit{Note: * indicates statistically significant difference (p < 0.05). WM = World Model, OV = Offline Videos.}
\end{table}