### 2025-06-16 results/fixed_rl_2025-06-13_19-28-06/logs/publication_paper/tables/main_results_real.tex
add runs

\begin{table*}[htbp]
\centering
\caption{Comprehensive Comparison of Learning Paradigms for Surgical Action Prediction}
\label{tab:main_results}
\begin{threeparttable}
\begin{tabular}{lcccccc}
\toprule
\textbf{Learning Paradigm} & \textbf{mAP\tnote{1}} & \textbf{Exact Match} & \textbf{Planning} & \textbf{Training} & \textbf{Inference} & \textbf{Paradigm} \\
                           & \textbf{(Mean ± Std)} & \textbf{Accuracy} & \textbf{Stability} & \textbf{Time (min)} & \textbf{Speed (fps)} & \textbf{Category} \\
\midrule
Supervised IL & 0.737 ± 0.020 & 0.328 & 0.998 & 2.1 & 145 & Supervised \\
RL + Direct Video\\(A2C) & 0.706 ± 0.023 & 0.320 & 1.000 & 12.1 & 102 & Model-Free RL \\
RL + Direct Video\\(PPO) & 0.705 ± 0.025 & 0.310 & 1.000 & 12.1 & 102 & Model-Free RL \\
RL + World Model\\(PPO) & 0.702 ± 0.022 & 0.300 & 1.000 & 14.3 & 98 & Model-Based RL \\
RL + World Model\\(A2C) & 0.701 ± 0.021 & 0.295 & 1.000 & 14.3 & 98 & Model-Based RL \\

\bottomrule
\end{tabular}
\begin{tablenotes}
\item[1] Mean Average Precision across all action classes. Statistical significance tests show p < 0.05 between Supervised IL and RL approaches.
\end{tablenotes}
\end{threeparttable}
\end{table*}