### 2025-07-05 results/2025-07-05_21-17-49/fold0/miccai_paper/component_analysis_table.tex

\begin{table}[h]
\centering
\caption{Component-wise Performance Analysis (IL Baseline)}
\label{tab:component_analysis}
\begin{tabular}{lccc}
\toprule
\textbf{Component} & \textbf{Current Recognition} & \textbf{Next Prediction} & \textbf{Degradation} \\
\midrule
Instrument (I) & 90.3\% & 87.9\% & 2.4\% \\
Verb (V) & 68.1\% & 65.1\% & 3.0\% \\
Target (T) & 57.1\% & 56.1\% & 1.0\% \\
Instrument-Verb (IV) & 45.7\% & 42.6\% & 3.1\% \\
Instrument-Target (IT) & 53.2\% & 52.3\% & 0.9\% \\
\bottomrule
\end{tabular}
\end{table}