### 2025-07-06 results/2025-07-06_17-21-56/fold0/miccai_paper/main_results_table.tex
new plots

\begin{table}[h]
\centering
\caption{Comparative Performance of IL and RL Approaches for Surgical Action Planning}
\label{tab:main_results}
\begin{tabular}{lccccc}
\toprule
\textbf{Method} & \textbf{Current mAP} & \textbf{Next mAP} & \textbf{1s Planning} & \textbf{10s Planning} & \textbf{Degradation} \\
\midrule
\textbf{IL Baseline} & \textbf{45.6\%} & \textbf{44.9\%} & \textbf{47.1\%} & \textbf{29.1\%} & \textbf{38.2\%} \\
IRL Enhanced & 44.2\% & 43.8\% & 45.3\% & 28.7\% & 36.6\% \\
World Model RL & 42.1\% & 41.6\% & 43.8\% & 27.2\% & 37.9\% \\
Direct Video RL & 43.9\% & 43.1\% & 44.9\% & 28.1\% & 37.4\% \\
\bottomrule
\end{tabular}
\end{table}